Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-08-26 paper.Rnw
proof reading edits through the description of the sine data in Section 3

% Set up for ASA data science journal
\articletype{Article Type}
\received{26 April 2016}
\revised{6 June 2016}
\accepted{6 June 2016}
\raggedbottom

@@

% Abstract and keywords
\abstract[Summary]{The importance of providing explanations for predictions made by black-box models has led to the development of model explainer methods such as LIME (local interpretable model-agnostic explanations) \citep{ribeiro:2016}. LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplot, (2) assessment metric plot, and (3) feature heatmap. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameters and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made about LIME \citep{ribeiro:2016} that are similar to other criticisms in the literature (\citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}).}
\keywords{LIME, black-box models, interpretability, diagnostics}

@@

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models are often used for the latter purpose. These models have proven to perform well in a wide range of prediction problems, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as health care and forensics science. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

@@

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, mohseni:2018, molnar:2019}. Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{welling:2016}), and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on \hh{one specific} model-agnostic method: LIME \citet{ribeiro:2016} \hh{introduced by Ribeiro, Singh, and Guestrin in 2016}.

%\subsection{Conceptual Description of LIME}

@@

@
\caption{A conceptual depiction of a faithful LIME explainer model in the immediate neighborhood of a prediction of interest. The predictions from a hypothetical black-box model are plotted against the standardized values of the two hypothetical predictor variables. The diamond shaped points represent the location of a prediction of interest. The size and opacity of the circular points indicate the weight assigned based on the distance to the prediction of interest computed using the inverse of the Gower distance metric raised to the power of \Sexpr{good_gower_power}. The black lines are a weighted ridge regression model used as an explainer model that reasonably captures the relationship between the black-box predictions and the features in a local region around the prediction of interest. That is, the explainer is faithful to the complex model and produces a reasonable explanation that Feature 1 plays a more important role in the prediction of interest than Feature 2 since the magnitude of the slope associated with Feature 1 is larger.}
\label{fig:concept-plot-good}

@@

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A ridge regression model is fit to the same hypothetical black-box model predictions and predictor variables as in \autoref{fig:concept-plot-good}, but the weights are computed using a Gower exponent of 1. The explainer model is unfaithful to the complex model in the immediate neighborhood of the prediction of interest.}
\label{fig:concept-plot-bad}
\end{figure*}

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two predictor variables used to fit the hypothetical model. The diamond shaped points represent a prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. \hh{In the example of \autoref{fig:concept-plot-good}}, an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of \hh{two features in the data} % features Features 1 and~2 
as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

The plot on the left of \autoref{fig:concept-plot-good} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of the prediction of interest. The explainer model captures the relationship in the immediate neighborhood around the prediction of interest  with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows no global or local relationship between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set as the observed value of the prediction of interest, and it has an appropriately small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.

%\subsection{Motivation for Diagnosing LIME}

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME\footnote{https://github.com/marcotcr/lime} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameters selected need to be assessed.

We consider an example where the choice of exponent for the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:concept-plot-bad} show the same data as \autoref{fig:concept-plot-good}, but the Gower distance metric exponent is decreased to \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package). This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:concept-plot-good}, the explainer model captures the relationship between the black-box predictions in the immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:concept-plot-bad}. In addition, the magnitude of the slope (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) associated with Feature 2 is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in predicting the observation of interest.

Several sources in the literature discuss the performance of LIME. One of the biggest difficulties with LIME is determining how to specify a local region \citep{laugel:2018} \citep{molnar:2019}. This is due to an unclear definition of a "local region" and how to apply LIME to achieve an appropriate local region as demonstrated by \autoref{fig:concept-plot-bad}. \citet{alvarezmelis:2018} raise a concern pertaining to the robustness of explanations from LIME and other model explainers: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, \citet{ribeiro:2016} acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state ``if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

@@

%\subsection{Structure of Paper}

The remainder of the paper is structured as follows. Section \ref{background} provides background and claims made by \citet{ribeiro:2016} about LIME. We introduce the suggested diagnostic plots in Section \ref{diagnostics}. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest model fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extensions and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}. The diagnostic plots included in this paper are created using the R package \emph{limeaid} \citep{goode:2020}. %\kgc{(Would you like me to add you as an author on the limeaid package?)}\hh{I appreciated your offer, but limeaid is your package - I haven't written a single line :)}

@@

LIME was introduced in 2016 by \citet{ribeiro:2016}. The  authors provide an implementation of LIME in a Python package\footnote{https://github.com/marcotcr/lime}. An adaption of the Python package in R has been implemented and made available by Thomas Lin-Pedersen \citep{pedersen:2020}. The discussion of parameter choices and implementation details of LIME in this paper are based on the R package.

@@

The assumption of interpretability only depends on the complexity of the model used as explainer model. If the model is too complex to provide meaningful explanations (e.g. there are too many variables in the model), it is clear that the assumption of interpretability is violated. The other three assumptions are not as easy to assess -- for those we suggest the use of diagnostic plots.

@@

# Separte the data into training and testing parts
set.seed(20191003)
rs <- sample(1:600, 500, replace = FALSE)
sine_data_train <- sine_data[rs,]
sine_data_test <- sine_data[-rs,]

@@

# Create points represting the rotated since curve
sinefun_data <- data.frame(xnew = seq(min(sine_data$x1new),
                                      max(sine_data$x1new),
                                      by = 0.01)) %>%
  mutate(ynew = 5 * sin(xnew)) %>%
  mutate(x = rot_x(xnew, ynew, -theta),
         y = rot_y(xnew, ynew, -theta)) %>%
  filter(y >= -10, y <= 10)

@@

@
\caption{Plots of $x_2$ versus $x_1$ from the training (left) and testing (right) sets of the \data \ introduced in Section \ref{diagnostics}. The true classification boundary is shown as the solid black line in both plots. The color of the training data points represents the value of the observed response variable ($y$). The color of the testing data points represents the random forest probability that an observation belongs to the category of blue ($\hat{y}$). The \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} cases that are misclassified by the random forest are identified by black circles. The prediction of interest to explain is indicated by a diamond.}
\label{fig:sine-plot-data}
\end{figure*}

@@

To demonstrate the visualizations, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with three features and one response variable. The features, $x_1$, $x_2$, and $x_3$, are randomly sampled from Unif(\Sexpr{min}, \Sexpr{max}), Unif(\Sexpr{min}, \Sexpr{max}), and $\mbox{N}(0,1)$ distributions, respectively. A binary response variable $y$ is  created using a rotated sine curve. In particular, let $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$ and $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$ where $\theta=-0.9$. Then $y$ is defined as
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ . %\nolabel
  \end{cases}
\end{eqnarray}
Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$.
\hh{From the plots in \autoref{sine-plot-data}}\kge{,} \hh{we see that the global relationship between response $y$ and features $x_1$ and $x_2$ is linear: the probability for label blue increases with the difference between features $x_2$ and $x_1$. Locally, the relationship between $y$ and features $x_2$ and $x_1$ varies a lot more around the line of identity. Here, the relationship is determined by the sine waves. However, the sine is a good-natured function that can be approximated well linearly} \kge{in local regions}.

The dataset is divided into a training set of \Sexpr{dim(sine_data_train)[1]} observations and a testing set of \Sexpr{dim(sine_data_test)[1]} observations. A random forest model is fit using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings. The model is applied to the test set to obtain predictions. \autoref{fig:sine-plot-data} shows scatterplots of $x_2$ versus $x_1$ from the training data (left) and the testing data (right). Both plots include the true classification boundary of the rotated sine function plotted as the solid black line. The training data are colored by the observed response variable ($y$), and the testing data are colored by the prediction probabilities from the random forest model ($\hat{y}$). The random forest model misclassifies \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} points on the classification boundary. These are identified by circles in \autoref{fig:sine-plot-data}.

For the presentation of the explanation scatterplot, we focus on the misclassified point with $(x_1, x_2, x_3)$ coordinates of (\Sexpr{round(sine_poi$x1, 2)}, \Sexpr{round(sine_poi$x2, 2)}, \Sexpr{round(sine_poi$x3, 2)}) indicated by a diamond in \autoref{fig:sine-plot-data}. Misclassified points are often of interest to explain since they may provide information about ways to improve the model. For the introduction of the other three plots, we use all observations in the \data test data as points of interest.

\hh{All runs of LIME in this paper are executed in}
a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. 
The forked version \hh{is functionally indistinguishable from Pedersen's implementation}\kge{,} \hh{but it} allows us to export internal values \hh{relevant for an assessment of the explainer}. 
The R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid} (\Sexpr{packageVersion("limeaid")}) implements diagnostic plots based on the output of the altered version of \emph{lime}.

@@

A quantile bin based simulation method is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, \hh{because the complexity of the explanations increases with  the number of bins.} The sixth application of LIME uses a kernel density simulation method. The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications.

\subsection{A LIME explanation}
\hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.} \kgc{Okay.} To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package. The figure shows that the random forest returned a probability of \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} that the observation of interest belongs to category blue {(denoted as the ``label" in the figure).The lengths of the bars represent the explainer model coefficients associated with the indicator variables chosen by LIME as important. The color of the bar denotes the sign of the coefficient and whether the feature supports or contradicts a random forest classification of `blue'. The "explanation fit" value is the deviance ratio from the R package \emph{glmnet} \citep{simon:2011} for a ridge regression model. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014}, and it should not be used as the only metric in a model assessment.

@@

The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest model. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist. It is difficult to assess linearity from an explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix \ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter} and shows a violation of the linearity claim.

>>>>

% Set up for ASA data science journal
\articletype{\kge{Research Article}}
\received{\kge{XX August 2020}}
\revised{\kge{XX XX XXXX}}
\accepted{\kge{XX XX XXXX}}
\raggedbottom

@@

% Abstract and keywords
\abstract[Summary]{The importance of providing explanations for predictions made by black-box models has led to the development of model explainer methods such as LIME (local interpretable model-agnostic explanations) \citep{ribeiro:2016}. LIME uses a surrogate model to explain the relationship between predictor variables and predictions from a black-box model in a local region around a prediction of interest. However, the quality of the resulting explanations relies on how well the explainer model captures the black-box model in a specified local region. Here we introduce three visual diagnostics to assess the quality of LIME explanations: (1) explanation scatterplot\kge{s}, (2) assessment metric plot\kge{s}, and (3) feature heatmap\kge{s}. We apply the visual diagnostics to a forensics bullet matching dataset to show examples where LIME explanations depend on the tuning parameters and the explainer model oversimplifies the black-box model. Our examples raise concerns about claims made \kge{of \sout{about}} LIME \citep{ribeiro:2016} that are similar to other criticisms in the literature (\citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}).}
\keywords{LIME, black-box models, interpretability, diagnostics}

@@

In the field of statistics, there are two main uses for models: inference and prediction. Machine learning models are often used for the latter purpose. These models have proven to perform well in a wide range of prediction problems, but the accuracy of many machine learning models comes at the cost of interpretability due to their algorithmic complexity (hence the phrase "black-box models"). Model interpretability allows for the understanding and assessment of how a model produces predictions. The lack of the ability to understand and assess a model makes it difficult to trust the model, especially in areas with high stakes decisions such as \kge{medical \sout{health care}} and forensics science\kge{s}. The increased use of machine learning models in applications and the introduction of the General Data Protection Regulation (GDPR) in 2018 \citep{goodman:2016} has resulted in a dramatic increase in explainable machine learning research, which focuses on developing ways to explain output from machine learning algorithms.

@@

Numerous methods have been proposed to provide explanations for black-box model predictions \citep{gilpin:2018, guidotti:2018, ming:2017, molnar:2019}. \kgc{I just took out the Mohseni reference since it is not peer reviewed. I think the other four should give enough of an idea of places to look for overviews on EML. Does that seem okay to you?} Some are specific to one type of model (e.g. \citep{simonyan:2013} and \citep{urbanek:2008}), \kgc{I switched to a reference to Urbanek's visualizations since the Welling paper was not peer reviewed. Is this better?} %\citep{welling:2016}), 
and others are model-agnostic (e.g. \citep{fisher:2018} and \citep{strumbelj:2014}). In this paper, we focus on \hh{one specific} model-agnostic method: LIME \citep{ribeiro:2016} \kge{(}\hh{introduced by Ribeiro, Singh, and Guestrin in 2016}\kge{)}. \kgc{Do we need the addition of when the paper was introduced here since the statement is made at the beginning of Section 2, or did you intend to remove that sentence?}

@@

@
\caption{A conceptual depiction of a faithful LIME explainer model in the immediate neighborhood of a prediction of interest. The predictions from a hypothetical black-box model are plotted against the standardized values of the two hypothetical predictor variables. The diamond shaped points represent the location of a prediction of interest. The size and opacity of the circular points indicate the weight assigned based on the distance to the prediction of interest computed using the inverse of the Gower distance metric raised to the power of \Sexpr{good_gower_power}. The black lines \kge{represent \sout{are}} a weighted ridge regression model used as an explainer model that reasonably captures the relationship between the black-box predictions and the features in a local region around the prediction of interest. That is, the explainer is faithful to the complex model and produces a reasonable explanation that Feature 1 plays a more important role in the prediction of interest than Feature 2 since the magnitude of the slope associated with Feature 1 is larger.}
\label{fig:concept-plot-good}

@@

@
\caption{A conceptual depiction of an unfaithful local explainer model in the immediate neighborhood of a prediction of interest. A ridge regression model is fit to the same hypothetical black-box model predictions and predictor variables as in \autoref{fig:concept-plot-good}, but the weights are computed using a Gower exponent of 1. The explainer model is unfaithful to the complex model in \kge{any \sout{the}} immediate neighborhood of the prediction of interest.}
\label{fig:concept-plot-bad}
\end{figure*}

\autoref{fig:concept-plot-good} provides a visualization of this conceptual understanding of LIME. The two plots show the predictions from a hypothetical black-box model plotted against the two \kge{hypothetical model} predictor variables \kge{\sout{used to fit the hypothetical model}}. The diamond shaped points represent \kge{\sout{a} the location of the} prediction of interest. A Gower distance metric \citep{gower:1971} is used to define locality such that the size of the points represent the inverse of the Gower distance raised to some value and indicate the proximity to the prediction of interest. \hh{In the example of \autoref{fig:concept-plot-good}}, an exponent of \Sexpr{good_gower_power} is used to emphasize a very local region around the prediction of interest. A ridge regression model weighted by the proximity values is used as the explainer model with the black-box predictions as the response variable and standardized versions of \hh{two features in the data} as predictor variables. Standardized features allow direct comparisons of the model coefficients. The explainer model is depicted by the black lines in the figure.

The plot on the left of \autoref{fig:concept-plot-good} shows the relationship between the black-box predictions and Feature 1. Here, the explainer model is plotted with Feature 2 fixed at the observed value of the prediction of interest. The explainer model captures the relationship in \kge{an \sout{the}} immediate neighborhood around the prediction of interest with a slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 1") %>% pull(slope) %>% round(3)}. The plot on the right shows no global or local relationship\kge{s} between the black-box predictions and Feature 2. Here, the explainer model is plotted with Feature 1 set as the observed value of the prediction of interest, and it has an appropriately small slope of \Sexpr{coefs_data %>% filter(case == "good", feature == "Feature 2") %>% pull(slope) %>% round(3)}. The magnitude of the slope associated with Feature 1 is larger than the slope of Feature 2, which suggests that Feature 1 plays a more important role in the prediction made by the black-box model for the prediction of interest. This explanation agrees with a visual assessment of the relationships between the predictions and predictor variables.

The concept of LIME is relatively simple: use an interpretable model to approximate a complex model in a local region. However, a practical implementation of LIME is not straightforward, and research is being done to improve the procedure \citep{laugel:2018}. The current implementations of LIME\footnote{https://github.com/marcotcr/lime}\kgc{Is there a better way to cite a Python package?} \citep{pedersen:2020} offer various tuning parameters (see Section \ref{background}) that affect the explainer model and ultimately, the explanation. Since the explainer model is an approximation of the complex model and not a direct interpretation, the explanations produced by an explainer model are subject to the quality of the approximation. Thus, in order to achieve accurate explanations, the tuning parameters selected need to be assessed.

We consider an example where the choice of exponent for the weights is crucial to the quality of the explanation. The plots in Figure \ref{fig:concept-plot-bad} show the same data as \autoref{fig:concept-plot-good}, but the Gower distance metric exponent is decreased to \Sexpr{bad_gower_power} (the default exponent in the \emph{lime} R package). This causes the observations that are further away from the prediction of interest to be given larger weights than before. In \autoref{fig:concept-plot-good}, the explainer model captures the relationship between the black-box predictions in \kge{an \sout{the}} immediate neighborhood of the prediction of interest. This cannot be said of the explainer model in \autoref{fig:concept-plot-bad}. In addition, the magnitude of the slope associated with Feature 2  (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 2") %>% pull(slope) %>% round(3) %>% abs()}) is larger than that of Feature 1 (\Sexpr{coefs_data %>% filter(case == "bad", feature == "Feature 1") %>% pull(slope) %>% round(3) %>% abs()}). Thus, this explainer model actually provides a misleading explanation that Feature 2 plays a more important role in \kge{the} predict\kge{ion\sout{ing the observation}} of interest.

Several sources in the literature discuss the performance of LIME. One of the biggest difficulties with LIME is determining how to specify a local region \citep{laugel:2018} \citep{molnar:2019}. This is due to an unclear definition of a "local region" and how to apply LIME to achieve an appropriate local region as demonstrated by \autoref{fig:concept-plot-bad}. \citet{alvarezmelis:2018} raise a concern pertaining to the robustness of explanations from LIME and other model explainers: they find that even small changes in predictor variables can lead to very different LIME explanations. Additionally, \citet{ribeiro:2016} acknowledge that if a linear model is used as the explainer, LIME relies on a linear approximation of the explainer model to the complex model and state "if the underlying model is highly non-linear even in the locality of the prediction, there may not be a faithful explanation".

@@

The remainder of the paper is structured as follows. Section \ref{background} provides background and claims made by \citet{ribeiro:2016} about LIME. We introduce the suggested diagnostic plots in Section \ref{diagnostics}. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest \kge{\sout{model}} fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extensions and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}. \kgc{I moved the following two sentences here from Section 2, so that it refers to the whole paper more clearly.} \hh{All runs of LIME in this paper are executed in} a forked version\footnote{https://github.com/goodekat/lime} of the development version of the R package \emph{lime} (\Sexpr{packageVersion("lime")}) \citep{pedersen:2020}. The forked version \hh{is functionally indistinguishable from Pedersen's implementation}\kge{,} \hh{but it} allows us to export internal values \hh{relevant for an assessment of the explainer}. %The R package \emph{limeaid} \footnote{Available on GitHub at https://github.com/goodekat/limeaid} (\Sexpr{packageVersion("limeaid")}) implements diagnostic plots based on the output of the altered version of \emph{lime}.
The diagnostic plots included in this paper are created using the R package \emph{limeaid} \citep{goode:2020}.

@@

LIME was introduced in 2016 by \citet{ribeiro:2016}. \kgc{You added this sentence to Section 1. Do you want it removed from here or there? If it is removed here, the next sentence will need to be changed.} The authors provide an implementation of LIME in a Python package\footnote{https://github.com/marcotcr/lime}. \kgc{(Again, is there a better way to cite a Python package?)} An adaption of the Python package in R has been implemented and made available by Thomas Lin-Pedersen \citep{pedersen:2020}. The discussion of parameter choices and implementation details of LIME in this paper are based on the R package.

@@

The assumption of interpretability only depends on the complexity of the model used as explainer model. If the model is too complex to provide meaningful explanations (e.g. there are too many variables in the model), it is clear that the assumption of interpretability is violated. The other three assumptions are not as easy to assess\kge{, and} %-- 
for those\kge{,} we suggest the use of diagnostic plots.

@@

# Separate the data into training and testing parts
set.seed(20191003)
rs <- sample(1:600, 500, replace = FALSE)
sine_data_train <- sine_data[rs,]
sine_data_test <- sine_data[-rs,]

@@

# Create points representing the rotated since curve
sinefun_data <- data.frame(xnew = seq(min(sine_data$x1new),
                                      max(sine_data$x1new),
                                      by = 0.01)) %>%
  mutate(ynew = 5 * sin(xnew)) %>%
  mutate(x = rot_x(xnew, ynew, -theta),
         y = rot_y(xnew, ynew, -theta)) %>%
  filter(y >= -10, y <= 10)

@@

@
\caption{Plots of $x_2$ versus $x_1$ from the training (left) and testing (right) sets of the \data \ introduced in Section \ref{diagnostics}. The true classification boundary is shown as the solid black line in both plots. The color of the training data \kge{\sout{points}} represents the value of the observed response variable ($y$). The color of the testing data \kge{\sout{points}} represents the random forest probability that an observation belongs to the category of blue \kge{\sout{($\hat{y}$)}}. The \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} cases that are misclassified by the random forest are identified by black circles. The prediction of interest to explain is indicated by a diamond.}
\label{fig:sine-plot-data}
\end{figure*}

@@

To demonstrate the \kge{visual diagnostics \sout{visualizations}}, we generate an example dataset that will be referred to as the \data. The \data \ contains \Sexpr{dim(sine_data)[1]} observations with three features and one response variable. The features, $x_1$, $x_2$, and $x_3$, are randomly sampled from Unif(\Sexpr{min}, \Sexpr{max}), Unif(\Sexpr{min}, \Sexpr{max}), and $\mbox{N}(0,1)$ distributions, respectively. A binary response variable $y$ is  created using a rotated sine curve. In particular, let $x'_1=x_1\cos(\theta)-x_2\sin(\theta)$ and $x'_2=x_1\sin(\theta)+x_2\cos(\theta)$ where $\theta=-0.9$. Then $y$ is defined as
\begin{eqnarray}\label{eq:data}
  y=\begin{cases}
  \mbox{blue} & \mbox{ if } x'_2 > \sin\left(x'_1\right) \\
  \mbox{red} & \mbox{ if } x'_2 \le \sin\left(x'_1\right) \ . %\nolabel
  \end{cases}
\end{eqnarray}
Note that due to the creation of $y$ in this manner, $y$ is dependent on $x_1$ and $x_2$ and independent of $x_3$. The dataset is divided into \kge{training and testing sets of \Sexpr{dim(sine_data_train)[1]} and \Sexpr{dim(sine_data_test)[1]} observations, respectively. \sout{a training set of \Sexpr{dim(sine_data_train)[1]} observations and a testing set of \Sexpr{dim(sine_data_test)[1]} observations.}} A random forest \kge{\sout{model}} is fit using the R package \emph{randomForest} (version \Sexpr{packageVersion("randomForest")}) \citep{liaw:2002} with the default settings. The model is applied to the test set to obtain predictions.

\autoref{fig:sine-plot-data} shows scatterplots of $x_2$ versus $x_1$ from the training data (left) and the testing data (right). Both plots include the true classification boundary of the rotated sine function plotted as the solid black line. The training data are colored by the observed response variable ($y$), and the testing data are colored by \kge{random forest \sout{the}} prediction probabilities \kge{\sout{from the random forest model ($\hat{y}$)}}. The random forest \kge{\sout{model}} misclassifies \Sexpr{dim(sine_data_test %>% filter(y != rfpred))[1]} points\kge{, which are all located near \sout{on}} the classification boundary. These are identified by \kge{open} circles in \autoref{fig:sine-plot-data}.

\kgc{I moved this paragraph. Figure 3 had not been introduced yet, so I think it fits better here.} \hh{From \kge{\sout{the plots in}} \autoref{fig:sine-plot-data}}\kge{,} \hh{we see that the global relationship between response $y$ and features $x_1$ and $x_2$ is linear: the probability for label blue increases with the difference between features $x_2$ and $x_1$. Locally, the relationship between $y$ and features $x_2$ and $x_1$ varies a lot more around the line of identity. Here, the relationship is determined by the sine waves. However, the sine is a good-natured function that can be approximated well linearly} \kge{in local regions}.

\kgc{I moved some paragraphs around here. It didn't read smoothly to me. I also added a sentence similar to one I had written before and you removed that I had included as an exposition sentence. Without it, I think it's confusing to jump into listing the tuning parameters we use without an overview of what is done. See what you think of it now.} \kge{We apply LIME using six sets of tuning parameters to all observations in the \data \ test set to observe variability across tuning parameters.} A quantile bin based simulation method is used for five of the LIME applications with the number of bins varying from 2 to 6 by application. We use 6 bins as the maximum, \hh{because the complexity of the explanations increases with the number of bins.} The sixth application of LIME uses a kernel density simulation method. The default methods for feature selection (forward selection) and the computation of the weights (Gower distance raised to an exponent of 1) are used for all applications. \kge{(See Appendix~\ref{lime-details} for more details on the tuning parameters.)}

For the presentation of the explanation scatterplot, we focus on the misclassified point with $(x_1, x_2, x_3)$ coordinates of (\Sexpr{round(sine_poi$x1, 2)}, \Sexpr{round(sine_poi$x2, 2)}, \Sexpr{round(sine_poi$x3, 2)}) indicated by a diamond in \autoref{fig:sine-plot-data}. Misclassified points are often of interest to explain since they may provide information about ways to improve the model. For the introduction of the other \kge{two \sout{three}} plots, we \kge{consider the LIME explanations for \sout{use}} all observations in the \data \ test data. \kge{\sout{as points of interest.}} \kgc{XXX This is where I left off XXX}

@@

%\subsection{A LIME explanation}

\paragraph{An Example LIME Explanation}

\kgc{I really like the idea of separating the description of the \emph{lime} figure from the explanation scatterplot, but the more I think about it, I don't like that it get it's own sub-section. I think it would be nicer to allocate the sub-sections to the visual diagnostic plots. Instead, I switched to using a heading like you created for introducing the sine data. I also changed the wording of the heading a bit. What do you think? Another option would be to drop the heading altogether and just say, "Here is an example of a LIME explanation...but it is not enough..."}

\hh{XXX I've just pulled the description of the left hand side of figure 4 out of the next section.} To demonstrate a LIME explanation, we focus on the misclassified point indicated in Figure \ref{fig:sine-plot-data} by the diamond as the prediction of interest. The left plot in Figure \ref{fig:sine-exp-scatter} shows a visualization of the explanation for the prediction of interest as suggested by Ribeiro et al and implemented in the \emph{lime} R package. The figure shows that the random forest returned a probability of \Sexpr{round(sine_poi_lime_default$label_prob[1], 2)} that the observation of interest belongs to category blue {(denoted as the ``label" in the figure).The lengths of the bars represent the explainer model coefficients associated with the indicator variables chosen by LIME as important. The color of the bar denotes the sign of the coefficient and whether the feature supports or contradicts a random forest classification of `blue'. The "explanation fit" value is the deviance ratio from the R package \emph{glmnet} \citep{simon:2011} for a ridge regression model. In other words, this is the $R^2$ value associated with the explainer model. In this case, it is \Sexpr{round(sine_poi_lime_default$model_r2[1], 2)} suggesting that the explainer model is not a good linear fit. However, it is commonly accepted that $R^2$ has limitations for assessing the quality of fit of a model \citep{sapra:2014}, and it should not be used as the only metric in a model assessment.

@@

The explanation scatterplot shows issues with a definition of locality and an oversimplification of the random forest \kge{\sout{model}}. LIME does provide an explanation for the prediction of interest, which makes sense based on the quantile bins used. However, the explanation scatterplot identifies that better explanations for the prediction of interest exist. It is difficult to assess linearity from an explanation scatterplot, but a residual plot of the explainer model could be used to check the linearity claim. See Appendix \ref{residual-plot} for the residual plot associated with the explanation considered in Figure \ref{fig:sine-exp-scatter} and shows a violation of the linearity claim.
